### 2021-07-15 plos_latex_template.tex
adding some of Andrea's comments to the text and changing the style

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

@@

\section*{Rule 1: Embrace all dimensions of diversity}
The first step to be inclusive is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event. Make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. Write a diversity statement (e.g. \url{https://user2021.r-project.org/about/diversity_statement/}) and express this welcoming spirit in social media to let people know that they are seen, respected and welcome. That this is their space and their community too. 
%ast be aware that participants may come from different backgrounds and that freedom of speech can be used to harm others, so don't let one dimension of diversity (i.e., political opinion) prevail over others. See rule 10.

\section*{Rule 2: Have an inclusive organizing team}
An inclusive conference can only be organized by an inclusive and diverse organizing team. Invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity. Pay special attention to the usually underrepresented groups. If the organization is not used to working with a diverse group, it may be challenging. But the challenges will help your team grow, learn and have more inclusive practices. If the organizing team is global and remote, you may struggle with time zones, and meet at odd times. You may also have to review carefully the choice of words for emails,
%ast ¿algo más enfático? [do the effort to] teach yourself the correct choice of words...
social media publications and your website, to be respectful of every culture and situation;
%ast yo cortaría esto en dos separadas. 
team members could be quick to spot text that is not right, or that does not resonate with a particular group of people. An inclusive team will let you know if the organization is failing in any aspect of inclusiveness or accessibility.

@@

When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, always look for the best in their subject. If there is not much diversity in your first set of names, it does not mean that minorities are not good enough. It probably means that the system has always privileged some groups of people, and that you need to look further to find great people that are not on the spotlight. [At least in one book I've read a discussion about H-index that we could cite here]. Look beyond your narrow and often limited networks. And if you start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.

@@

Conferences, even virtual ones, should have a registration fee mainly for two reasons. The first one is that preparing the conference takes a lot of effort and costs money (e.g. commercial registration tools, captioning, or conference venue if in-person) and it should be reflected on a price. The second reason is that people tend to value more the things they pay for, and there is a higher risk of not showing up to a free even [any publication to back this up?]. On the other hand, if you are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees. Usually, there is a higher fee for people from the industry than for academia. Consider also a lower fee for non-profit organizations, government employees, or free lancers. 
% ast: don't forget people in vulnerable job positions, such as post-docs
You can also use conversion factors for country of residence using data from International Comparison Program report of the World Bank.

@@

Offering scholarships or grants to attend your conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g. their PI could pay for it), but could apply for it to enrich their CV. If the goal is to help the people in need, make that clear. Do not give it a fancy name. It is a fee waiver or a discount for the people in need. And simplify the process for applicants. People with low resources already have a hard time applying for loans, grants and scholarships in their lives. Do not make life harder on them by asking them to write long paragraphs to convince you that they deserve your support. Many people that cannot afford conferences do not even try to apply for waivers/scholarships because they think it would be hard to get them. Trust that when they say they need it, it is because they do.

If your resources allow it, you could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).

@@

Conferences are usually inaccessible to people with hearing and visual impairments. Take them into account in every stage of the process. Welcome them to your organizing team so that they can be part of the decisions since the beginning, and remind you of inaccessible practices that need to improve. Communicate on social media with alternative text and videos with captions. Use screen-reader friendly platforms for conference registration, abstract submission and, if the conference is virtual or hybrid, look for a screen-reader friendly chat platform too. Provide captioning for presentations. If the conference is in-person, make sure that presenters speak to the microphone to make it easier for captioners or interpreters to listen to them. Provide accessibility guidelines for slides and presentations, and encourage the presenters to follow them. Accessibility means, for example, that the speakers should also provide raw and accessible material to their talks, such as RMarkdown, html or tex files, and that, if they pre-record their presentation, to make sure that their video is shown and their face is visible so that people with hearing impairment can have the possibility to read their lips. You can also give accessibility awards at the end of the conference to acknowledge the speakers who were mindful of inclusiveness when preparing their talks. Please find an example of accessibility guidelines at \url{https://user2021.r-project.org/participation/accessibility/}.
%ast Bear in mind that most accessibility practices are beneficial to everyone. Captions, for instance, are used by non-native speakers and people that access the conference in loud environments. The prior availability of material benefits people with low bandwidth and non-native speakers as well.

\section*{Rule 7: Put more effort in promoting your conference among groups of people who are usually underrepresented}

If some minorities are under represented in your conference, they may not feel part of the community or may not have the financial resources to attend. Promote the conference in both ways: as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
For the former, if part of the community has been historically discriminated, emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them. For the latter, advertise the fee waivers, financial support, and the easiness of the process. People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be rather supportive in this case.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials and workshops are in English. While one official language is extremely helpful to communicate and English is the primary language in scientific communication, opening the conference to other languages could make it less intimidating to people who are not so fluent in English. Excluding them may also mean missing great innovative pieces of work, because of a language barrier. So advertise your conference in several languages, and consider having workshops and presentations in other languages (with or without captions in English).

@@

To maintain and inclusive and safe space in your conference, there should be a code of conduct and a team to enforce it. `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces. The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' (coc book)

The code of conduct team should receive training to have a clear understanding on what to enforce and what not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate the responses. A diverse code of conduct team could be more understanding of intersectionality issues in discrimination and harassment practices. 
%the code of condcut should be displayed proeminently in several spaces of the conference and it should be effective if it deters people from unaccepted behavior in the first place

>>>>

\section*{Introduction}

@@

\section*{Rule 1: Embrace all dimensions of diversity}
The first step to be inclusive is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event. We can make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. Publishing a diversity statement (e.g. \url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in social media will let people know that they are seen, respected and welcome. That this is their space and their community too. 
%ast be aware that participants may come from different backgrounds and that freedom of speech can be used to harm others, so don't let one dimension of diversity (i.e., political opinion) prevail over others. See rule 10.

\section*{Rule 2: Have an inclusive organizing team}
An inclusive conference can only be organized by an inclusive and diverse organizing team. We can invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity. Special attention should be payed to the usually underrepresented groups. If the organization is not used to working with a diverse group, it may be challenging. But the challenges will help the team grow, learn and have more inclusive practices. If the organizing team is global and remote, time zones can be hard to manage, and meetings will end up happening at odd times. The team may also have to review carefully and learn to choose the right words for emails,
social media publications and your website, to be respectful of every culture and situation. 
Team members could be quick to spot text that is not right, or that does not resonate with a particular group of people. An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it.

@@

When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, we should always look for the best in their subject. If there is not much diversity in the first set of names, it does not mean that minorities are not good enough. It probably means that the system has always privileged some groups of people, and that we need to look further to find great people that are not on the spotlight. [At least in one book I've read a discussion about H-index that we could cite here]. Let's look beyond our narrow and often limited networks. And if we start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.

@@

Conferences, even virtual ones, should have a registration fee mainly for two reasons. The first one is that preparing the conference takes a lot of effort and costs money (e.g. commercial registration tools, captioning, or conference venue if in-person) and it should be reflected on a price. The second reason is that people tend to value more the things they pay for, and there is a higher risk of not showing up to a free even [any publication to back this up?]. On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees. Usually, there is a higher fee for people from the industry than for academia. We should also consider a lower fee for non-profit organizations, government employees, or free lancers. And use conversion factors for country of residence using data from International Comparison Program report of the World Bank. It is important to include discounts for students as well as for postdocs, as these are often precarious positions. Postdoc status is not always well-defined in academia and can vary for each country, so their payment category should be explicitly defined.

@@

Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g. their PI could pay for it), but could apply for it to enrich their CV. If the goal is to help the people in need, let's make that clear. Do not give it a fancy name. It is a fee waiver or a discount for the people in need. And the process for applicants should be simplified. People with low resources already have a hard time applying for loans, grants and scholarships in their lives. We should not make their lives harder by asking them to write long paragraphs to convince us that they deserve our support. Many people that cannot afford conferences do not even try to apply for waivers/scholarships because they think it would be hard to get them. We should trust that when they say they need it, it is because they do.

If the conference resources allow it, one could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).

@@

Conferences are usually inaccessible to people with hearing and visual impairments. They should be taken into account in every stage of the process. Welcoming them into the organizing team would allow them to take part in the decisions since the beginning, and remind us of inaccessible practices that need to improve. In social media communication, images should be accompanied with alternative text and videos with captions. Platforms for conference registration, abstract submission --and, if the conference is virtual or hybrid, chat platforms --should be screen-reader friendly (tested beforehand). Captioning for presentations should be available. If the conference is in-person, presenters should speak to the microphone to make it easier for captioners or interpreters to listen to them. Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged. Accessibility means, for example, that the speakers should also provide raw and accessible material to their talks, such as RMarkdown, html or tex files, and that, if they pre-record their presentation, to make sure that their video is shown and their face is visible so that people with hearing impairment can have the possibility to read their lips. Accessibility awards at the end of the conference are a good way to acknowledge the speakers who were mindful of inclusiveness when preparing their talks. Please find an example of accessibility guidelines at \url{https://user2021.r-project.org/participation/accessibility/}.

Accessibility practices should not be left for the last minute. They require time and early decision-making. Conversely, inaccessible decisions are hard to undo; e.g. when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
Bear in mind that most accessibility practices are beneficial to everyone. Captions, for instance, are used by non-native speakers and people that access the conference in loud environments. The prior availability of material benefits people with low bandwidth and non-native speakers as well.

\section*{Rule 7: Put more effort in promoting the conference among groups of people who are usually underrepresented}

If some minorities are under represented in the conference, they may not feel part of the community or may not have the financial resources to attend. And effort must be made to promote the conference as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
For the former, if part of the community has been historically discriminated, one should emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them. For the latter, fee waivers, financial support, and the easiness of the process should be advertised. People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be rather supportive in this case.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials and workshops are in English. While one official language is extremely helpful to communicate and English is the primary language in scientific communication, opening the conference to other languages could make it less intimidating to people who are not so fluent in English. Excluding them may also mean missing great innovative pieces of work, because of a language barrier. Advertising the conference in several languages, and considering having workshops and presentations in other languages (with or without captions in English) could help overcome this barrier.

@@

To maintain and inclusive and safe space in the conference, there should be a code of conduct and a team to enforce it. `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces. The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' (coc book)
The code of conduct should be displayed prominently in several spaces of the conference to deter people from incurring in unaccepted behavior.

The code of conduct team should receive training to have a clear understanding on what to enforce and what not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate the responses. A diverse code of conduct team could be more understanding of intersectionality issues in discrimination and harassment practices.
